### 2024-04-12 outputs/paper/paper.qmd
Update paper.qmd

---
title: "Are we living in a depressive society?"
subtitle: "Trends in symptoms of depressive disorder in the US population from May 2020 to March 2024"
author: Chay Park
thanks: "Code and data are available at: https://github.com/Chay-HyunminPark/Anxiety_depressive_disorder."
date: today
date-format: long
abstract: "This study explores how socio-cultural factors like age, sex, race/ethnicity, and education level influenced depressive disorder in the U.S. during and after the COVID-19 pandemic, analyzing data from 2020 to 2024. Using advanced statistical methods in R, it uncovered significant variations in mental health impacts across different demographic groups. The research reveals a complex relationship between socio-cultural factors and mental health, highlighting the increased vulnerability of certain populations during global crises. The findings contribute to the development of targeted public health measures to alleviate mental health issues amid and following pandemics, enhancing our understanding of the impact of socio-cultural factors on mental well-being in times of crisis."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

# Load necessary libraries for the paper
library(tidyverse)
library(lubridate)
library(arrow)
library(here)
library(ggrepel)  # For better label placement
library(extrafont)
library(ggplot2)
library(gridExtra)

```

# Introduction

Depression and anxiety are not only prevalent but escalating issues in the modern world, affecting individuals across the globe regardless of demographic factors (World Health Organization, 2021). These mental health challenges have been further exacerbated by the COVID-19 pandemic, which led to widespread social and economic upheaval (Brooks et al., 2020). While various studies have explored the impact of the pandemic on mental health, there remains a significant gap in understanding the interactions between socio-cultural factors and the prevalence of anxiety and depressive disorders during this period (Pfefferbaum & North, 2020).

This paper addresses this gap by analyzing data from the U.S. Census Bureau's Household Pulse Survey, initiated in response to the pandemic to assess its effects on American households, including mental wellness (U.S. Census Bureau, 2020). Our analysis focuses on the correlation between the prevalence of anxiety or depressive disorders and socio-cultural factors such as age, sex, race/ethnicity, and education level during the pandemic and post pandemic years of 2020 to 2024. Specifically, the estimand of this study is the difference in the rates of symptoms of anxiety or depressive disorders across individuals, correlated with socio-cultural factors including age, sex, race/ethnicity, and education level.

Employing a quantitative methodology, this paper utilizes R (R Core Team 2020) for analysis. We utilized packages like ggplot2 (Wickham et al., 2016) and gridExtra (Murrell, 2021) for data analysis and visualization. Our findings reveal significant trends and disparities in the rates of anxiety and depressive disorders across different socio-cultural groups, highlighting the intricate relationship between these factors and mental health outcomes during the pandemic.

The importance of this research lies in its potential to inform public health policies and interventions, aiming to mitigate the adverse effects of such global crises on mental health (Xiong et al., 2020). By understanding the specific socio-cultural factors that contribute to increased vulnerability to anxiety and depressive disorders, targeted strategies can be developed to support affected populations more effectively.

The paper is structured to first present the trend analysis of anxiety and depressive disorder rates in the U.S. from 2020 to 2024, followed by an examination of how these rates vary across different demographic groups. Subsequent sections discuss the statistical methods and results for each research question. It opens up the findings and let the world explores their broader implications for mental health research and policy. The conclusion summarizes the key insights, acknowledges the limitations of the study, and suggests avenues for future research.

The study focuses on answering the following research questions:

* What is the trend in US anxiety and depressive disorder rates from 2020 to 2024?

* How do depressive rates vary across different age groups?

* How do depressive rates differ based on race and ethnicity?

* How do depressive rates differ based on sex?

* How do adepressive rates differ based on education level?

* Is there codependency of social connection on one's depression level?

Through this analysis, we aim to better understand contributing factors of mental wellness of the US population. Furthermore, we display possible correlation between factors and the mentall wellness states.

# Data {#sec-data}

## Methodology

The U.S. Census Bureau, in collaboration with five federal agencies, launched the Household Pulse Survey to produce data on the social and economic impacts of Covid-19 on American households. The Household Pulse Survey was designed to gauge the impact of the pandemic on employment status, consumer spending, food security, housing, education disruptions, and dimensions of physical and mental wellness (National Center for Health Statistics, 2024). In the paper, we will focus on the impact of the pandemic on mental wellness by analysing the dataset of the indicators of anxiety or depression based on reported frequency of symptoms. The data on the trend in US anxiety and depressive disorder rates, measured from 2020 to 2024, was sourced from the Centers for Disease Control and Prevention (CDC). Alongside with 'National Estimate' group in which covers the whole population trends, trends in anxiety and depressive disorder rates by population subgroup, specifically categorized by age, race and ethnicity, sex, education groups were gathered from the National Center for Health Statistics (NCHS). The data is freely available at https://data.cdc.gov/NCHS/Indicators-of-Anxiety-or-Depression-Based-on-Repor/8pt5-q6wp/about_data, with raw files located in the inputs/data on the GitHub repository.

## Features

In this paper, data validation is conducted using an R script file named '03-test_data' to ensure the accuracy of the dataset obtained directly from the source. The paper utilizes both 'Indicators_of_Anxiety_or_Depression_Based_on_Reported_Frequency_of_Symptoms_During_
Last_7_Days_20240401' and 'Lack_of_Social_Connection' CSV files. Then by the characteristics of the group, the data sets are separately saved as 'Age_subgroup_trends' where it only saves the rows with 'By Age' under the 'Group' column, irrespective of race and ethnicity. Similarly, 'Educ_subgroup_trends' are the dataset only with the rows 'By Education' under the 'Group' column to identify the patterns of each population subgroup with classification of highest educational attainment, regardless of race and ethnicity. All CSV files are saved as parquet file. Due to its functionality, as opposed to row-based formats like CSV or Excel, Parquet allows it to efficiently handle analytical queries by accessing only the necessary columns of data, which reduces storage reads, improves query performance, and lowers storage costs. The column 'Value' is used to denote the symptom rate on the y-axis. 
Column 'Time Period End Date' is used to denote the range of the period, a date for the x-axis.

# Results

## Trend in US Anxiety and Depressive Disorder Rates

The @fig-andtrend-graph overall trend in US anxiety and depressive disorder rates fluctuate over time, with the highest rates recorded in the end of 2020 and the beginning of 2021 at an average rate above 40% symptom rates for both anxiety disorder or depressive disorder. Subsequently, symptoms of anxiety disorder followed by the second showing similar trends as the first. Lastly, symptoms of depressive disorder show the relativelys lower rates. + add more.

## Trend in US Anxiety and Depressive Disorder Rates

The overall trend in US anxiety and depressive disorder rates shows significant fluctuations, with several notable peaks indicating heightened rates of reported symptoms. For instance, one of the highest peaks occurs in early 2021, where the rate of symptoms for combined anxiety and depressive disorders reaches approximately 45%. This peak reflects a substantial increase compared to earlier periods. Another significant peak, observed in mid-2022, sees rates approaching closer to 36-37%, illustrating another period of notable increase in symptoms among the population.

These peaks are interspersed with valleys, such as in late 2020, where rates temporarily dip to around 25%, showing the episodic nature of these conditions. The plot also highlights a general trend of increasing rates from the onset of the data in 2020 through to peaks in 2021, followed by some fluctuation but maintaining relatively high levels overall.

The data indicates that the combined symptom rates of both anxiety and depressive disorders generally present higher values, consistently staying above 30% after the initial 2020 data points. This trend underscores the overlapping nature of these mental health issues and their persistent impact on the US population.

The graph, depicted as @fig-andtrend-graph, provides a line graph visualization of these rates over time. This visual representation allows for clear observation of trends and seasonal variations, with specific dates marked on the x-axis to aid in pinpointing changes. For example, a significant increase is noted as we move from late 2020 into early 2021, correlating with various global and national events that may have influenced mental health.

## Trends in Depressive Disorder Rates by Population Subgroup by Age

This analysis highlights the variations in depressive disorder rates across different age subgroups in the US, utilizing data up to early 2024. The visualization provided in the graph titled @fig-ageandtrend-graph reveals distinct temporal trends across various age demographics, effectively illustrating how depressive symptoms fluctuate over time among these groups.

### Data Handling and Visualization Techniques
The data is sourced from a Parquet file, ensuring robust and efficient handling of large datasets. For accuracy in temporal analyses, the 'Date' column is meticulously converted to a date format. The script focuses on depressive symptoms by age subgroup, filtering the data to showcase how these rates vary across different age demographics. A line graph, enhanced with different colors for each age group, facilitates easy tracking of each subgroup's symptom rates over time.

### Annotation and Aesthetic Customization
On January 10, 2022, the graph incorporates annotations for each age group, pinpointing the rates of depressive symptoms at this specific date. These annotations are strategically placed using `geom_text_repel` to prevent overlap, ensuring clarity and readability. The graph is further refined with horizontal lines at regular 10% intervals from 0 to 50%, providing a scale to assess the symptom rates. Custom dates along the x-axis highlight key moments in the timeline, aiding in the analysis of changes over time.

### Thematic and Stylistic Elements
The plot employs a classic theme with tailored font settings, which not only enhances the visual appeal but also ensures the readability of textual elements. The absence of a legend, dictated by the plot's customization, focuses attention on the lines themselves, each representing a different age group’s trend.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-andtrend-graph
#| fig-cap: "Trend in US Anxiety and Depressive Disorder Rates"

# Load the fonts
loadfonts(device = "win")

# Define the path to the data file and load it
data <- read_parquet(file = here::here("outputs/data/Overall_trends.parquet"))

# Filter the dataset for the specified symptoms
filtered_data <- data %>%
  filter(Indicator %in% c('Symptoms of Depressive Disorder', 
                          'Symptoms of Anxiety Disorder', 
                          'Symptoms of Anxiety Disorder or Depressive Disorder')) %>%
  mutate(Indicator = factor(Indicator, levels = c('Symptoms of Anxiety Disorder or Depressive Disorder',
                                                  'Symptoms of Anxiety Disorder',
                                                  'Symptoms of Depressive Disorder')))

# Ensure that 'Time.Period.End.Date' is a Date object
filtered_data$Date <- as.Date(filtered_data$Date, format = "%Y-%m-%d")

# Define the specific dates for the x-axis scale
specific_dates <- as.Date(c("2020-05-05", "2020-09-14", "2021-01-05", 
                            "2021-05-10", "2021-09-13", "2022-01-10", 
                            "2022-05-09", "2022-09-13", "2023-01-16",
                            "2023-05-08", "2023-09-04", "2024-01-08"))

# Plotting the line graph
p <- ggplot(filtered_data, aes(x = Date, y = Value, group = Indicator, color = Indicator)) +
  geom_line() +
  geom_hline(yintercept = seq(10, 45, by = 5), color = "lightgrey") +
  scale_x_date(breaks = specific_dates, date_labels = "%Y-%m") +
  labs(x = NULL,
       y = "Symptom Rates (%)") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, family = "Times"),
        axis.text.y = element_text(family = "Times"),
        axis.title.x = element_text(family = "Times"),
        axis.title.y = element_text(family = "Times"),
        legend.position = "bottom",
        legend.box = "vertical",
        legend.direction = "vertical",
        plot.margin = unit(c(1, 1, 1, 1), "cm"))

# Display the plot
print(p)

```

## Trends in Depressive Disorder Rates by Population Subgroup

### Population subgroup by age

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-ageandtrend-graph
#| fig-cap: "Depressive Disorder Trend by Population Subgroup by Age"

# Load necessary libraries
library(tidyverse)
library(arrow)
library(here)
library(ggrepel)
library(extrafont)

# Load the fonts
loadfonts(device = "win")

# Define the path to the data file and load it
data <- read_parquet(file = here::here("outputs/data/Age_subgroup_trends.parquet"))

# Convert 'Date' column to Date format if it's not already
data$Date <- as.Date(data$Date, format = "%Y-%m-%d")

# Define the specific dates for the x-axis scale
specific_dates <- as.Date(c("2020-05-05", "2020-09-14", "2021-01-05", 
                            "2021-05-10", "2021-09-13", "2022-01-10", 
                            "2022-05-09", "2022-09-13", "2023-01-16",
                            "2023-05-08", "2023-09-04", "2024-01-08"))

# Filter the data to get the rows for the specific date to label
label_date <- as.Date("2022-01-10")
label_data <- data %>%
  filter(Date == label_date)

# Plotting the line graph
p <- ggplot(data, aes(x = Date, y = Value, group = Subgroup, color = Subgroup)) +
  geom_line() +
  geom_text_repel(data = label_data, aes(label = Subgroup, x = Date, y = Value), 
                  family = "Times", color = "black",
                  size = 3,  # Smaller text size
                  arrow = arrow(type = "closed", length = unit(0.02, "inches"), angle = 15),
                  box.padding = unit(0.5, "lines"),
                  point.padding = unit(0.5, "lines")) +
  geom_hline(yintercept = seq(0, 50, by = 10), color = "lightgrey") +
  scale_x_date(breaks = specific_dates, date_labels = "%Y-%m") +
  labs(x = NULL, y = "Symptom Rates (%)") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, family = "Times"),
        axis.text.y = element_text(family = "Times"),
        axis.title.x = element_text(family = "Times"),
        axis.title.y = element_text(family = "Times"),
        legend.position = "none",
        plot.margin = unit(c(1, 1, 1, 1), "cm"))

# Display the plot
print(p)

```

### Population subgroup by sex

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-sexandtrend-graph
#| fig-cap: "Depressive Disorder Trend by Population Subgroup by Sex"

# Define the path to the data file and load it
data <- read_parquet(file = here::here("outputs/data/Sex_subgroup_trends.parquet"))

# Convert 'Date' column to Date format if it's not already
data$Date <- as.Date(data$Date, format = "%Y-%m-%d")

# Define the specific dates for the x-axis scale
specific_dates <- as.Date(c("2020-05-05", "2020-09-14", "2021-01-05", 
                            "2021-05-10", "2021-09-13", "2022-01-10", 
                            "2022-05-09", "2022-09-13", "2023-01-16",
                            "2023-05-08", "2023-09-04", "2024-01-08"))

# Filter the data to get the rows for the specific date to label
label_date <- as.Date("2022-01-10")
label_data <- data %>%
  filter(Date == label_date)

# Plotting the line graph
p <- ggplot(data, aes(x = Date, y = Value, group = Subgroup, color = Subgroup)) +
  geom_line() +
  geom_hline(yintercept = seq(0, 40, by = 5), color = "lightgrey") +
  scale_x_date(breaks = specific_dates, date_labels = "%Y-%m") +
  labs(x = NULL, y = "Symptom Rates (%)") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, family = "Times"),
        axis.text.y = element_text(family = "Times"),
        axis.title.x = element_text(family = "Times"),
        axis.title.y = element_text(family = "Times"),
        legend.position = "none",  # Remove the separate legend
        plot.margin = unit(c(1, 1, 1, 1), "cm"))

p + geom_text_repel(
  data = label_data,
  aes(label = Subgroup),
  nudge_x = 0,  # Adjust nudge_x and nudge_y to correctly position the labels
  nudge_y = 0,
  direction = "y",
  color = "black",
  family = "Times"
)

```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-raceandtrend-graph
#| fig-cap: "Depressive Disorder Trend by Population Subgroup by Race"

# Load necessary libraries
library(tidyverse)
library(arrow)
library(here)
library(ggrepel)
library(extrafont)

# Load the fonts
loadfonts(device = "win")

# Define the path to the data file and load it
data <- read_parquet(file = here::here("outputs/data/Race_subgroup_trends.parquet"))

# Convert 'Date' column to Date format if it's not already
data$Date <- as.Date(data$Date, format = "%Y-%m-%d")

# Rename the subgroup values
data$Subgroup <- recode(data$Subgroup,
                        "Hispanic or Latino" = "Hispanic",
                        "Non-Hispanic Asian, single race" = "Asian",
                        "Non-Hispanic Black, single race" = "African American",
                        "Non-Hispanic White, single race" = "Caucasian",
                        "Non-Hispanic, other races and multiple races" = "Other")

# Define the specific dates for the x-axis scale
specific_dates <- as.Date(c("2020-05-05", "2020-09-14", "2021-01-05", 
                            "2021-05-10", "2021-09-13", "2022-01-10", 
                            "2022-05-09", "2022-09-13", "2023-01-16",
                            "2023-05-08", "2023-09-04", "2024-01-08"))

# Filter the data to get the rows for the specific date to label
label_date <- as.Date("2022-01-10")
label_data <- data %>%
  filter(Date == label_date)

# Plotting the line graph
p <- ggplot(data, aes(x = Date, y = Value, group = Subgroup, color = Subgroup)) +
  geom_line() +
  geom_text_repel(data = label_data, aes(label = Subgroup, x = Date, y = Value), 
                  nudge_x = 0, nudge_y = 0, family = "Times", color = "black",
                  size = 3,  # Make label text smaller
                  arrow = arrow(type = "closed", length = unit(0.02, "inches"), ends = "last", angle = 30)) +
  geom_hline(yintercept = seq(0, 40, by = 5), color = "lightgrey") +
  scale_x_date(breaks = specific_dates, date_labels = "%Y-%m") +
  labs(x = NULL, y = "Symptom Rates (%)") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, family = "Times"),
        axis.text.y = element_text(family = "Times"),
        axis.title.x = element_text(family = "Times"),
        axis.title.y = element_text(family = "Times"),
        legend.position = "none",
        plot.margin = unit(c(1, 1, 1, 1), "cm"))

# Display the plot
print(p)

```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-educandtrend-graph
#| fig-cap: "Depressive Disorder Trend by Population Subgroup by Education"

library(tidyverse)
library(arrow)
library(here)
library(ggrepel)
library(extrafont)

# Load the fonts
loadfonts(device = "win")

# Define the path to the data file and load it
data <- read_parquet(file = here::here("outputs/data/Educ_subgroup_trends.parquet"))

# Convert 'Date' column to Date format if it's not already
data$Date <- as.Date(data$Date, format = "%Y-%m-%d")

# Rename the subgroup values
data$Subgroup <- recode(data$Subgroup,
                        "Bachelor's degree or higher" = "College graduate",
                        "High school diploma or GED" = "High school graduate",
                        "Some college/Associate's degree" = "Some college",
                        "Less than a high school diploma" = "No high school degree")

# Define the specific dates for the x-axis scale
specific_dates <- as.Date(c("2020-05-05", "2020-09-14", "2021-01-05", 
                            "2021-05-10", "2021-09-13", "2022-01-10", 
                            "2022-05-09", "2022-09-13", "2023-01-16",
                            "2023-05-08", "2023-09-04", "2024-01-08"))

# Prepare data for labels
label_data <- data %>%
  group_by(Subgroup) %>%
  summarize(Date = median(Date), Value = median(Value)) %>%
  ungroup()

# Plotting the line graph with adjusted label positions
p <- ggplot(data, aes(x = Date, y = Value, group = Subgroup, color = Subgroup)) +
  geom_line() +
  geom_text_repel(data = label_data, aes(label = Subgroup, x = Date, y = Value), 
                  family = "Times", 
                  size = 3,  # Smaller text size
                  arrow = arrow(type = "closed", length = unit(0.02, "inches"), angle = 30),
                  box.padding = unit(0.5, "lines"),
                  point.padding = unit(0.5, "lines")) +
  geom_hline(yintercept = seq(0, 40, by = 5), color = "lightgrey") +
  scale_x_date(breaks = specific_dates, date_labels = "%Y-%m") +
  labs(x = "Date", y = "Symptom Rates (%)") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, family = "Times"),
        axis.text.y = element_text(family = "Times"),
        axis.title.x = element_text(family = "Times"),
        axis.title.y = element_text(family = "Times"),
        legend.position = "none",
        plot.margin = unit(c(1, 1, 1, 1), "cm"))

print(p)

```
```{r}

# Define the path to the data file and load it
social_data <- read_parquet(file = here::here("outputs/data/Social_Connection.parquet"))

# Filter data for the indicator "Adults who usually or always feel lonely" and "By Age" subgroup
social_age_data <- social_data %>%
  filter(Indicator == "Adults who usually or always feel lonely", Group == "By Age")

# Plot the stacked bar chart
graph_age <- ggplot(social_age_data, aes(x = `Time.Period`, y = Value, fill = Subgroup)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = "Adults who usually or always feel lonely - By Age",
       x = NULL,
       y = "Value") +
  theme_minimal()

# Filter data for the indicator "Adults who usually or always feel lonely" and "By Sex" subgroup
social_sex_data <- social_data %>%
  filter(Indicator == "Adults who usually or always feel lonely", Group == "By Sex")

# Plot the stacked bar chart
graph_sex <- ggplot(social_sex_data, aes(x = `Time.Period`, y = Value, fill = Subgroup)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = "Adults who usually or always feel lonely - By Sex",
       x = NULL,
       y = "Value") +
  theme_minimal()

# Filter data for the indicator "Adults who usually or always feel lonely" and "By Race/Hispanic ethnicity" subgroup
social_race_data <- social_data %>%
  filter(Indicator == "Adults who usually or always feel lonely", Group == "By Race/Hispanic ethnicity")

# Plot the stacked bar chart
graph_race <- ggplot(social_race_data, aes(x = `Time.Period`, y = Value, fill = Subgroup)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = "Adults who usually or always feel lonely - By Race",
       x = NULL,
       y = "Value") +
  theme_minimal()

# Filter data for the indicator "Adults who usually or always feel lonely" and "By Education" subgroup
social_educ_data <- social_data %>%
  filter(Indicator == "Adults who usually or always feel lonely", Group == "By Education")

# Plot the stacked bar chart
graph_educ <- ggplot(social_educ_data, aes(x = `Time.Period`, y = Value, fill = Subgroup)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = "Adults who usually or always feel lonely - By Education",
       x = NULL,
       y = "Value") +
  theme_minimal()

# Arrange graphs in single pane
grid.arrange(graph_age, graph_sex, graph_race, graph_educ, nrow = 2, ncol = 2)

```

# Model

## Model set-up

### Model justification

# Discussion

What is done in this paper? 
What is something that we learn about the world? 
What is another thing that we learn about the world?

## Findings

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

Both Anxiety and Depressive Disorder showing the similar trends over the period. From this, we can tell either the number of participants ~

## Ethical Implication

## Limitation

-- What are some weaknesses of what was done? -- 
The survey was designed to meet the goal of accurate and timely weekly estimates. Hence, it was conducted by an internet questionnaire, with invitations to participate sent by email and text message (National Center for Health Statistics, 2024). Internet questionnaires face several significant challenges. One of the primary concerns is self-selection bias, where the survey tends to attract respondents who are particularly interested in the topic, which can skew the results and limit their generalizability. Additionally, these surveys may not reach a diverse demographic due to varying levels of internet access across different population segments, thus excluding certain groups, such as older adults or lower-income individuals, and potentially leading to a lack of representativeness. Respondents might also misunderstand or misinterpret questions without the opportunity for real-time clarification, compromising the accuracy of the data collected. Privacy and security concerns are prevalent too, as individuals may be hesitant to share personal or sensitive information online for fear of data breaches or doubts about the survey platform's security measures. Lastly, the validity and honesty of responses are at risk; while the anonymity of online surveys can encourage candidness, it might also lead to less thoughtful or dishonest answers, affecting the reliability of the survey outcomes (Dillman, Smyth, & Christian, 2014; Morling, 2018).

## Future Research
-- What is left to learn or how should we proceed in the future? --
For future research, survey can further ask what are the factors that affect their mood.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

## Diagnostics

\newpage

# References

---
nocite: '@*'
---